### 2018-08-11 CS761-RandomizedAlgorithms/assignments/final.tex
finished

%Problem 3
\newpage
\section{}
\paragraph{}
We throw $n^{0.99}$ balls into $n$ bins uniformly at random. Using union bound, the probability that a bin has at least $k$ balls, for $k$ chosen later, is at most 
$${n \choose k}(\frac{1}{n})^k.$$
Using the bound ${n^{0.99}\choose k} <(\frac{n^{0.99}e}{k})^k$ this probability is at most
$$(\frac{n^{0.99}e}{k})^k(\frac{1}{n})^k= \frac{e^k}{n^{0.0k}k^k}\leq \frac{e^k}{k^k}$$
where the last inequality follows since we will choose $k \geq 0$ and since $n\geq 1$ this implies $n^\frac{k}{100} \geq 1$.
Again invoking union bound
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq n\frac{e^k}{k^k} = n(\frac{e}{k})^k.$$ 
Since we want
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq \frac{1}{2}$$
we need to find $k$ such that
$$n(\frac{e}{k})^k \leq \frac{1}{2}.$$

%Problem 4
\newpage
\section{}
\paragraph{}
We always use addition, $+$, over the appropriate field (ie. $\F_n$ for inputs and $\F_m$ for outputs). Our algorithm is very simple. Given input $z$ choose $x \in \{0,\dots, n-1\}$ uniformly at random. Let $y = z-x$. Lookup $F(x)$ and $F(y)$ and return $F(x) + F(y)$.
\paragraph{}
If both $F(x)$ and $F(y)$ are not modified by our adversary then our return value satisfies
$$F(x) + F(y) = F(x+y) = F(z)$$
using the linearity of $F$ and the fact that $x+y = z$. So a necessary condition for our algorithm to fail is that one of $F(x)$ or $F(y)$ was modified by our adversary. For a uniform random element of the table, the probability of modification is at most $\frac{1}{5}$. Hence
$$Pr(F(x)\text{ or } F(y)\text{ was modified}) \leq Pr(F(x) \text{ was modified}) + Pr(F(y)\text{ was modified}) \leq \frac{2}{5}.$$
Therefore the probability of failure is at most
$$\frac{2}{5} < \frac{1}{2}.$$
Hence our algorithm succeeds with probability at least $\frac{1}{2}$ as desired. $\blacksquare$

@@

Let $\cA$ be an algorithm which obtain an $(\epsilon, \frac{1}{4})$-approximation to some parameter $\mu$. That is, when $X$ is the output value of $\cA$,

@@

>>>>

%Problem 3
\newpage
\section{}
\paragraph{}
We throw $n^{0.99}$ balls into $n$ bins uniformly at random.  The probability that a given bin has $k$ balls is 
$$p_k = {n^{0.99} \choose k} (\frac{1}{n})^k(1-\frac{1}{n})^{n^{0.99}-k} = \frac{1}{k!}\frac{n^{0.99}(n^{0.99}-1)\dots(n^{0.99} - k + 1)}{n^k}(1-\frac{1}{n})^{n^{0.99}-k}.$$
This value is approximately
$$p_k \approx \frac{e^{-n^{0.01}}(n^{0.01})^k}{k!}.$$
In other words, the probability that a given has load at least $k$ is a Poisson random variable.  In Poisson, all bins can be treated as independent so the probability that no bin has load at least $k$ is at most
$$(1-\frac{e^{-n^{0.01}}(n^{0.01})^k}{k!})^n\leq \exp(\frac{-e^{-n^{0.01}}n^{1 + \frac{k}{100}}}{k!})$$
To translate the Poisson case back to the exact case, Theorem $5.7$ tells us our final probability will be effected by a factor of at most $e\sqrt{n^{0.99}}$. Since we want a final probability of at most $\frac{1}{2}$ we need to choose $k$ so that 
$$\exp(\frac{-e^{-n^{0.01}}n^{1 + \frac{k}{100}}}{k!}) \leq \frac{1}{2e\sqrt{n^{0.99}}}$$
Or equivalently
$$\frac{-e^{-n^{0.01}}n^{1 + \frac{k}{100}}}{k!} \leq -\ln(2e\sqrt{n^{0.99}}).$$
Equivalently yet again,
$$k! \leq \frac{e^{-n^{0.01}}n^{1 + \frac{k}{100}}}{\ln(2e\sqrt{n^{0.99}})}$$
From $(5.5)$ of Probability and Computing, 
$$k! \leq k(\frac{k}{e})^k.$$
So again we can rewrite our desired inequality as
$$k(\frac{k}{e})^k \leq \frac{e^{-n^{0.01}}n^{1 + \frac{k}{100}}}{\ln(2e\sqrt{n^{0.99}})}$$
and taking logarithms
$$2k\ln(k) \leq \ln(e^{-n^{0.01}}n^{1 + \frac{k}{100}}) - \ln(\ln(2e\sqrt{n^{0.99}}) )= -n^{0.01} + (1+\frac{k}{100})\ln(n) - \ln(\ln(2e\sqrt{n^{0.99}})).$$
Now if we move the $k$ over from the right hand side and multiply the constants through we get the desired inequality
$$k(2\ln(k) - \frac{\ln(n)}{100}) \leq -n^{0.01} + \ln(n) - \ln(\ln(2e\sqrt{n^{0.99}})).$$
I ran out of time, but I hope that solving this equation for $k$ would yield a decent bound.
%Problem 4
\newpage
\section{}
\paragraph{}
We always use addition, $+$, over the appropriate field (ie. $\F_n$ for inputs and $\F_m$ for outputs). Our algorithm is very simple. Given input $z$ choose $x \in \{0,\dots, n-1\}$ uniformly at random. Let $y = z-x$. Lookup $F(x)$ and $F(y)$ and return $F(x) + F(y)$.
\paragraph{}
If both $F(x)$ and $F(y)$ are not modified by our adversary then our return value satisfies
$$F(x) + F(y) = F(x+y) = F(z)$$
using the linearity of $F$ and the fact that $x+y = z$. So a necessary condition for our algorithm to fail is that one of $F(x)$ or $F(y)$ was modified by our adversary. For a uniform random element of the table, the probability of modification is at most $\frac{1}{5}$. Hence
$$Pr(F(x)\text{ or } F(y)\text{ was modified}) \leq Pr(F(x) \text{ was modified}) + Pr(F(y)\text{ was modified}) \leq \frac{2}{5}.$$
Therefore the probability of failure is at most
$$\frac{2}{5} < \frac{1}{2}.$$
Hence our algorithm succeeds with probability at least $\frac{1}{2}$ as desired. $\blacksquare$

@@

Let $\cA$ be an algorithm which obtains an $(\epsilon, \frac{1}{4})$-approximation to some parameter $\mu$. That is, when $X$ is the output value of $\cA$,

@@

Sorry I didn't have time for this problem. My vague idea was to draw a connection between hitting times and the desired probabilities, then relate them back to $\phi$ on some electrical network.
